---
title: "Tidyverse - US Births"
output:
  html_document:
    keep_md: yes
    self_contained: no
  pdf_document: default
---
```{r, echo = FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="##", fig.retina=2, fig.path = "README_figs/README-")
```
#### Author 

- Joby John


#### Summary

In this  exercise, I am using US_births_2000-2014_SSA.csv dataset. I am using the dplyr functions to sort, group, mutate and summarize data.  

#### Load Libraries
```{r load-libraries, eval=TRUE, message=FALSE, warning=FALSE}

library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)

options(knitr.table.format = "html")
```




Following steps are done : 

- Read file into a dataframe

- Used dplyr function create summary and display in a kable

- USed dplyr to sort, group and mutate data

- Ploted number of births using ggplot



```{r load-data,message=FALSE, warning=FALSE }
data_url <-'https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_2000-2014_SSA.csv'
births_data <- read.csv(data_url)

births_data_summary_by_year <-    births_data %>%
                                  group_by( year )  %>% 
                                  group_by(year, total_births = sum(births)) %>%
                                  distinct(year,total_births)%>%
                                  select(year, total_births)%>%
                                  arrange(desc(total_births))
```

#### Number of births per year

```{r}
births_data_summary_by_year %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))  

#Births peaked in 2007. Any correlation to great recession?

ggplot(births_data_summary_by_year, aes(x= year, y= total_births)) + geom_line()

```

#### Number of births per month
```{r}
births_data_summary_by_month <-   births_data %>%
                                  mutate(month = month.abb[month])%>%
                                  group_by( month )  %>% 
                                  group_by(month, total_births = sum(births)) %>%
                                  distinct(month,total_births)%>%
                                  select(month, total_births)%>%
                                  arrange(desc(total_births))

births_data_summary_by_month %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))  

```

#### Top 10 days 
```{r}
top_10_birth_days<- births_data %>% 
            top_n(10, births) %>%
            select(year, month,date_of_month,births)%>%
            mutate(month = month.abb[month])%>%
            arrange(desc(births))

top_10_birth_days %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))  
```

#### Births per day
```{r}
births_data_summary_by_day_of_week <-    births_data %>%
                                  group_by(day_of_week )  %>% 
                                  group_by(day_of_week, total_births = sum(births)) %>%
                                  distinct(day_of_week,total_births)%>%
                                  select(day_of_week, total_births) %>%
                                  arrange(desc(total_births))
births_data_summary_by_day_of_week %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))  

```

#### Improvements

- Are Super Bowl Babies A Real Thing? We could Use this dataset to see if that claim is valid or not. 

- Convert numerical day of the week to actual day of the week. For example, 1 to sunday.

- Further analysis could be done. For example, how does economy affects the number of births.


#### References

 https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_2000-2014_SSA.csv

 https://www.bustle.com/articles/140278-are-super-bowl-babies-a-real-thing-one-commercial-makes-an-interesting-claim-video
 
